---
title: "pml_proj"
author: "Becky"
date: "Sunday, July 20, 2014"
output: html_document
---
#Synopsis
This study use the data from [http://groupware.les.inf.puc-rio.br/har] which collected Human Activity by machine and then aim to predict the activity into 5 classes (sitting-down, standing-up, standing, walking, and sitting).

In this paper, we will use the data mentioned above to develop a predictive model for this classes.


#Loading the dataset
```{r}
if(!file.exists("train.csv")){
  download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv","train.csv")
  }
train= read.csv("train.csv", na.strings=c("NA", "","#DIV/0!"))
```

#Processing the data
```{r}
dim(train)
summary(train$classe)
```
There are `r dim(train)[1]` observations with `r dim(train)[2]` variables. 

#Spliting the training and testing data
```{r}
library(caret)
set.seed(1234)
inTrain=createDataPartition(y=train$classe, p=0.7, list=FALSE)
t1=train[inTrain,]
t2=train[-inTrain,]

dim(t1)
```

#Selecting variables
```{r}
na_check= sapply(t1, function(x) {sum(is.na(x))})
table(na_check)
```
Only `r table(na_check)[1]` columns which contains less proportion of missing values, that can be used for further analysis.

```{r}
t11 = t1[, !names(t1) %in% names(na_check[na_check!=0])]
str(t11)
```
By looking at the summary of the data, we can see that the first 7 columns are recording the specification of records which useless to our model. We also take out these variables from the analysis.

```{r}
t11 = t11[,-c(1:7)]

```

#Building the model
We now use random forest method to create the prediction model. 

```{r, cache=TRUE}
ctrl <- trainControl(method = "repeatedcv", 
                     repeats = 5)
 
fit = train(classe~., data=t11, method="rf", preProcess=c("center","scale"),trControl=ctrl)
```

#Evaluating the model
After used the training dataset to develop the model, we use the testing dataset to evaluate the out-sample error rate.

```{r}
cf=confusionMatrix(t2$classe, predict(fit, t2))
```
The out-sample error rate is `r cf$overall[1]`.

#Reference: 
Data source [http://groupware.les.inf.puc-rio.br/har]

```{r include=FALSE}
submit=read.csv("test.csv", na.strings=c("NA", "","#DIV/0!"))
answers=predict(fit, submit)
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}
pml_write_files(answers)
```
